A Classical Density-Functional Theory for Describing Water Interfaces

### 2013-07-31 papers/water-saft/paper.tex
edit beginning of introduction Tested everything on node09.

Water \fixme{stuff about water}.  An important aspects of water
is its ability to form hydrogen bonds.  Each water molecule contains
two hydrogen bond donor sites and two acceptor sites.

One promising approach for an efficient continuum description of water
is that of classical density-functional theory (DFT), which is an
approach for evaluating the free energy and thermally averaged density
of fluids in an arbitrary external potential~\cite{ebner1976density}.
The foundation of classical DFT is the Mermin
theorem\cite{mermin1965thermal}, which extends the Hohenberg-Kohn
theorem\cite{hohenberg1964inhomogeneous} to non-zero temperature.
Classical DFT is a natural framework for creating a more flexible
theory of hydrophobicity that can readily describe interaction of
water with arbitrary external potentials---such as potentials
describing strong interactions with solutes or surfaces.

Statistical Associating Fluid Theory (SAFT) is a theory describing
complex fluids in which hydrogen bonding plays a significant
role\cite{chapman1989saft, muller2001molecular}. SAFT is used to
accurately model the equations of state of both pure fluids and
mixtures over a wide range of temperatures and pressures. In a recent
paper \cite{hughes2013classical}, we constructed a classical DFT for
water which uses SAFT as a basis for a free energy functional. We
used Clark \emph{et al}'s \cite{clark2006developing} five fitting
parameters for an optimal cDFT for water and, in addition, introduced
a parameter for the length scale over which the dispersion interaction
is correlated.

>>>>

Water, the universal solvent, is of critical practical importance, and
a continuum description of water is in high demand for a solvation
model.  A number of recent attempts to develop improved solvation
models for water have built on the approach of classical
density-functional theory (DFT)~\fixme{add citations here,
  particularly Jeanmairet 2013 and his citations 35-44, possibly
  cutting non-acqueous papers from that list}.
Classical DFT is based on a description of a fluid written as a free
energy functional of the density distribution.  There are two
approaches used to construct a classical DFT for water.  The first is
to choose a functional form that is fit to properties of the bulk
liquid at a given temperature and pressure~\fixme{cite Jeanmairet and
  similar}.  Using this approach, it is possible to construct a
functional that reproduces the exact second-order response function of
the liquid under the fitted conditions.  This class of functional,
however, will be less accurate at other temperatures or
pressures---and in the inhomogeneous scenarios in which solvation
models are applied.  The second approach is construct a functional by
applying liquid-state theory to a model system, and then fit the model
to experimental data such as the equation of state.

%% \fixme{stuff about water}.  An important aspects of water
%% is its ability to form hydrogen bonds.  Each water molecule contains
%% two hydrogen bond donor sites and two acceptor sites.
%% %
%% One promising approach for an efficient continuum description of water
%% is that of classical density-functional theory (DFT), which is an
%% approach for evaluating the free energy and thermally averaged density
%% of fluids in an arbitrary external potential~\cite{ebner1976density}.
%% The foundation of classical DFT is the Mermin
%% theorem\cite{mermin1965thermal}, which extends the Hohenberg-Kohn
%% theorem\cite{hohenberg1964inhomogeneous} to non-zero temperature.
%% Classical DFT is a natural framework for creating a more flexible
%% theory of hydrophobicity that can readily describe interaction of
%% water with arbitrary external potentials---such as potentials
%% describing strong interactions with solutes or surfaces.

A widely used family of models used in the development of classical
density functionals is based on Statistical Associating Fluid Theory
(SAFT).
%
SAFT is a theory for fluids in which hydrogen bonding plays a significant
role\cite{chapman1989saft, muller2001molecular}. SAFT is used to
accurately model the equations of state of both pure fluids and
mixtures over a wide range of temperatures and pressures.
\fixme{Talk about and cite papers using SAFT to build DFTs.}
%
In a recent paper \cite{hughes2013classical}, we constructed a
classical DFT for water which uses SAFT as a basis for a free energy
functional. We used Clark \emph{et al}'s \cite{clark2006developing}
five fitting parameters for an optimal cDFT for water and, in
addition, introduced a parameter for the length scale over which the
dispersion interaction is correlated.
